The Galactic Chemical Evolution of Carbon: Implications for Stellar Nucleosynthesis

### 2023-07-03 carbon_paper/tex/main.tex
removed uneccessary apendicies

\newcommand{\cc}{CCSNe}
\newcommand{\agb}{AGB}
\newcommand{\ia}{SNe Ia}
\newcommand{\sfh}{SFH}
\newcommand{\dla}{DLA}
\newcommand{\ssp}{SSP}
\newcommand{\imf}{IMF}
\newcommand{\gce}{GCE}
\newcommand{\apogee}{APOGEE}

\newcommand{\lims}{lower-intermediate mass stars}
\newcommand{\hms}{high-mass stars}

\defcitealias{james+21}{J21}
\newcommand{\citetjack}{Roberts et al.~(2023, in prep.)}
\newcommand{\citepjack}{(Roberts et al.~2023, in prep.)}
\newcommand{\citealtjack}{Robert et al.~2023, in prep.}

\newcommand{\cxi}{\texttt{\hyperlink{C11}{C11}}}
\newcommand{\kx}{\texttt{\hyperlink{K10}{K10}}}
\newcommand{\kxvi}{\texttt{\hyperlink{K16}{K16}}}
\newcommand{\vxiii}{\texttt{\hyperlink{V13}{V13}}}

\newcommand{\JJ}{\citetalias{james+21}}
\newcommand{\VICE}{\texttt{VICE}}
\newcommand{\caah}{[C/Mg]-[Mg/H]}
\newcommand{\caafe}{[C/Mg]-[Mg/Fe]}

\newcommand{\alp}{$\alpha$}
\newcommand{\Ycc}{\ensuremath{y_{\rm C}^{\rm cc}}}
\newcommand{\Yoc}{\ensuremath{y_{\rm Mg}^{\rm cc}}}
\newcommand{\Ycagb}{\ensuremath{y_{\rm C}^{\rm agb}}}
\newcommand{\Z}{Z}
\newcommand{\Zo}{%
    \ifmmode Z_{\sun}%
    \else $Z_{\sun}$%
    \fi}
\newcommand{\Mo}{%
    \ifmmode {\rm M}_{\sun}%
    \else M$_{\sun}$
    \fi}
\newcommand{\about}[1]{${\sim} #1$}

\newcommand{\proton}{\isotope[1][1]{p}}
\newcommand{\zoo}{\ensuremath{Z/Z_{\sun }}}
\newcommand{\Yct}{\ensuremath{y_{\rm C}}}

@@

To quantify yields, we define the stellar yield to be the fraction of a stars initial mass which is newly synthesized and released as a given element. 
For an element $X$ and star with mass $M$, the net-fractional stellar yield $\tilde{y}$ is 
\begin{equation}
\tilde{y}_{X} = \frac{M_{\rm ejected}}{M_{\rm birth}} (Z_{f, X} -  Z_{0, X})
\end{equation}
where $M_{\rm ejected}$ and $M_{\rm birth}$  are the total ejected mass and the birth mass of the star, and $Z_{f, X}$ and $Z_{0, X}$ are the initial birth abundance and final abundance of the ejected envelop respectably. For example, if a 1\,\Mo\ star has $\tilde{y}_{\rm C} = 10^{-3}$, then the star will add $10^{-3}\,\Mo$ of new C to the interstellar medium. 
Although per-star yields are necessary to compute \agb{} star enrichment rates in \gce{}  models, \imf-averaged yields are useful in interpreting their predictions. For a yield $y$ from a star of mass M and initial metallicity $Z$, the \imf-averaged yield is given by 
\begin{equation} \label{eq:imf-yield}
    y_{\rm X}(Z,t) = 
    \int_{M_{\rm min}(t)}^{M_{\rm max}} 
    \tilde{y}_{\rm X}(M, Z)
    \frac{dN}{dM}  \ dM
\end{equation}
where ${dN}/{dM}$ is the normalized \imf, $M_{\rm max}=100\,\Mo$ is the maximum stellar mass, and $M_{\rm min}(t)$ is the mass of stars with lifetime $t$.\footnotemark{}
We use $t=10\,$Gyr for total yields when $t$ is not used.
To calculate the \imf-averaged net yields, we use the Versatile Integrator for Chemical Evolution code (\VICE\footnotemark{}).

@@

	\begin{tabular}{l l l l}
		\hline
        Element & $\Ycc$ & $\tilde{y}^{\rm AGB}$ & $y^{\rm SNe Ia}$ \\
		\hline
        C & $0.0028 + 0.001\,(Z/Z_\odot)$ & $2.9\times{\rm C11}$ &  0 \\
        Mg & 0.00185 & 0 & 0 \\
        Fe & 0.0012 & 0 & 0.00214 \\
        N & 0.00072 & $9\times10^{-4}\,(Z/Z_\odot)\,M$ & 0\\
		\hline
	\end{tabular}
\end{table}

@@

An \agb\ star is a low-mass star ($\lesssim 8 M_{\sun}$) during its final phase of evolution.  
In an \agb\ star, two competing processes determine the outcome of C production: \textit{third dredge up} and \textit{hot bottom burning}.  
Third dredge up accompanies thermal pulses in \agb\ stars, where material from the CO core is mixed with the envelope, increasing surface C abundances \citep{KL14}. The C yields of the star are increased as this C-enhanced envelope is released to the interstellar medium. 
Hot bottom burning is the activation of the CNO cycle\footnotemark{}
at the bottom of the convective envelope when $T\gtrsim 50\,{\rm MK}$. Because the $^{14}$N proton capture is the slowest component of the CNO cycle, the CNO cycle converts nearly all \isotope[12]{C} into \isotope[14]{N} \citep{solar-fusion}.

Hot bottom burning and third dredge-up result in mass-dependent C yields. 
Stars less than about??2\,\Mo do not experience third dredge-up. As a result, these stars C abundances are only affected by first dredge-up, resulting in little change to C yields or slight destruction of C.
In 2?? to 4?? \Mo stars, third dredge up becomes important, enriching the outer layers with C, making these stars the most important producers of C. 
In massive, high-metallicity \agb\ stars, both hot bottom burning and third dredge up occur; however, third dredge up is much more efficient, resulting in significant \isotope[12]{C} destruction.

\footnotetext{
    The CNO cycle and the details of AGB evolution are described in appendix \ref{sec:stel_evo}.
}

    In this work, we explore four different sets of \agb\ star yield tables from literature,
    providing necessary well-sampled grids in mass and metallicity. We refer to the yields from the following studies as the following.
\begin{itemize}
    \item \cxi: \citet{cristallo+11, cristallo+15}
    \item \kx: \citet{karakas10}
    \item \vxiii: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \kxvi: \citet{KL16, karakas+18}
\end{itemize}
For our models to match observations, we find that need to uniformly amplify these yield tables. We use \cxi\ table, amplified by a factor of 2.9, as the fiducial \agb\ yield.

@@

Fig.~\ref{fig:agb-ssp}, on the left, shows the total production of C by \agb\ stars in a \ssp{} at an age $t$, i.e. $\tilde{y}_{\rm C}(Z_\odot, t)$. 
As the mass range $2\,\Mo\lesssim M \lesssim 4\,\Mo$ is most important for C production, about half of C production occurs before \about{1}\,Gyr, similar to \ia\ Fe. 
\kx{} and \kxvi{} weight C production more heavily towards high-mass AGB stars resulting in a faster enrichment delay time, whereas the \cxi\ and \vxiii\ models predict a slightly longer timescale of \about{1}\,Gyr. In any case, little to no C is produced more than 2\,Gyr after a star formation event. In contrast, Fe production continues steadily for 10\,Gyr.

@@

\begin{figure*}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}
        \caption[]{The net fractional \agb\ C yield  plotted as a function of initial stellar mass $M$ and colour-coded according to metallicity. The black dashed line shows $\tilde{y}=0$ for reference. Each panel represents yields from one of four \agb\ models: \cxi, \kx{}, \vxiii{}, \kxvi{} (see Section \ref{sec:agb} and  Appendix \ref{sec:oob_models}). }
        \label{fig:y_agb}
\end{figure*}

@@

Massive stars form $^{12}$C in their cores through the triple--$\alpha$ process (see Appendix \ref{sec:stel_evo}). However, only C ejected through supernovae and stellar winds contributes to the yield. 
While there are many stellar models providing predictions of \cc{} yields, the results of these models are highly uncertain due to the many stellar modeling uncertainties.

@@

\cc\ models do not reach [O/Mg] due to overproduction of O or underproduction of Mg, or both \citep{emily+21}. Here, we assume ${\rm [O/Mg]} = 0$, which is not compatible with \cc\ models but is consistent with \apogee{} observations \citep{weinberg+19, weinberg+22}.
Out model predictions of [C/Mg] and [C/H] are identical. There is a slight slope in [O/Mg] with metallicity, so this could add a systematic to our model.

@@

We assume a simple \textit{one-zone} chemical evolution model \cite[e.g.][]{tinsley80, pagel09, matteucci21}.  Newly produced metals are homogeneously and instantaneously mixed, so spatial dependence is neglected.
We define $M_{X}$ to be the mass of element $X$ in the gas-phase, $\dot{M}_\star$ to be the star formation rate (in $\Mo\,{\rm yr}^{-1}$), and $\eta$ to be the mass loading factor $\eta\equiv\dot{M}_{\rm outflow}/\dot{M}_\star$ (representing the strength of outflows). A \ssp\ returns a fraction $r$ of their birth mass to the interstellar medium, due to ejected stellar envelopes.\footnotemark{} 
Given the \imf-averaged yield of Mg $y_{\rm Mg}$, the rate of change in the gas-phase mass of Mg is a simple sum of sources and sinks,
\begin{equation}
    \dot{M}_{\rm Mg} =  y_{\rm Mg}\,\dot{M}_\star - \dot{M}_{\rm Mg,remnants} - \dot{M}_{\rm Mg,outflows},
\end{equation}
where $y_{\rm Mg} \dot{M}_{\star}$ describes \cc\ enrichment. 
In terms of the return mass fraction of stars $r$, the mass lost to remnants is $Z_{\rm Mg}\,\dot{M}_\star\,(1-r)$.  And, the outflows deplete mass at a rate $Z_{\rm Mg} \,\dot{M}_\star\,\eta$. (We assume the composition of outflows is the same as the interstellar medium.) Substituting for $\eta$ and $r$,  
\begin{equation}
    \dot{M}_{\rm Mg}= y_{Mg}\,\dot{M}_\star - (1 + \eta - r)\,Z_{Mg}\,\dot{M}_\star.
\end{equation}
Assuming an exponentially declining star formation history $\dot{M}_\star \propto e^{-t/\tau_{\rm sfh}}$, the equilibrium abundance is derived analytically by setting $\dot{Z}_{Mg}=0$.
\begin{equation}\label{eq:z_eq}
    Z_{\rm Mg}^{\rm eq}(R) = \frac{y_{\rm Mg}}{1 + \eta(R) - r - \tau_\star / \tau_{sfh}},
\end{equation}
where $\tau_{\star}$ is the star formation rate. ($\eta$ depends on $R$ to create a metallicity gradient. See also Section \ref{sec:vice} and Eq.~\ref{eq:eta_r}.)
In the special case of constant star formation, $\tau_{\rm sfh}\to\infty$, the denominator simplifies to $1+\eta-r$.

Likewise, for \agb\ stars,%
\footnote{In detail, the \agb\ contribution requires integration over the \sfh\ to account for the finite lifetimes of stars. However, for this exponential \sfh, this effect is minimal, so we simply use $\Ycagb$ for the current \agb\ contribution.}
Eq.~\ref{eq:z_eq} then becomes
\begin{equation}
    Z_{\rm C}^{\rm eq}(R) = \frac{\Ycc + \Ycagb}{1 + \eta(R) - r - \tau_\star / \tau_{\rm sfh}}
\end{equation}
And the equilibrium C/Mg abundance ratio is
\begin{equation}\label{eq:z_co}
    \frac{Z_{\rm C}^{\rm eq}}{Z_{\rm Mg}^{\rm eq}} = \frac{\Ycc + \Ycagb }{\Yoc}.
\end{equation}
Analogous to \cite{james+23} arguments about N, the trends in abundance ratios are set by yield ratios in these \gce{} models. The effect of other \gce{} parameters (most importantly $\eta$) cancels. As a consequence, yield ratios should establish abundance ratio trends in models which assume a different normalization of element yields and mass-loading (see discussion below).
This argument can also be inverted to infer  yields from abundance ratio trends. To the extent that observed C and Mg trends reflect the equilibrium abundances in different Galactic regions, we can infer the \cc\ yield given an assumed \agb\ star yield (or vice versa). Inferring $\Ycc$ from $\Ycagb$, 
\begin{equation}
    y_\text{C}^\text{CC} =  y_\text{Mg}^\text{CC} \frac{Z_\text{C,~eq}}{Z_\text{Mg,~eq}} - \langle y_c^{agb} \rangle
\end{equation}
Rewriting this expression as a relative yield of C and Mg,
\begin{equation}\label{eq:y_c_eq}
    \frac{\Ycc}{\Yoc} = \frac{Z_{\text{C},~\odot}}{Z_{\text{Mg},~\odot}} 10^{\rm [C/Mg]} - \frac{\Ycagb}{\Yoc}.
\end{equation}

@@

As we will discuss in Section \ref{sec:outflows}, the normalization of yields and $\eta$ is degenerate. This can be observed in Eq.~\ref{eq:z_co}, where changes in $\eta$ or the scaling of $y_{\rm C}/y_{\rm Mg}$ would not affect equilibrium trends. Furthermore, in Eq.~\ref{eq:z_eq}, an increase in both $y_{\rm Mg}$ and $\eta$ would leave $Z_{\rm Mg}^{\rm eq}$ unchanged. Our models here are unable to distinguish the overall scaling of yields and outflow mass loading. Choosing to keep $y_{\rm C}$ fixed reduces unnecessary free parameters. So, at solar metallicity, we set
\begin{subequations}
    \begin{align}
        \Yct~\big\vert_{Z=Z_\odot} &= \Ycc + \Ycagb = 0.005 \\
        \Yct/\Yoc~\big\vert_{Z=Z_\odot} &= 2.7.
    \end{align}
\end{subequations}
This ratio results in an equilibrium abundance ${\rm [C/Mg]} = -0.09$, which is consistent with the subgiant sample and is within \about{20\%} of the solar C/Mg mixture from \citet{asplund+09}.

In Section \ref{sec:f-z-models}, we will show that none of the four \agb\ yield sets (\cxi{}, \kx{}, \vxiii{}, \kxvi{}) produce enough C relative to this $\Ycc$ value. So, we introduce normalization factors, $\alpha_{\rm AGB}$, and $\alpha_{\rm CC}$ which denote a multiplicative scaling of $\Ycc$ and $\Ycagb$ respectively. 
\begin{subequations} \label{eq:alpha}
    \begin{align}
        \Ycagb &\rightarrow \alpha_{\rm AGB}\ \Ycagb \\
        \Ycc &\rightarrow \alpha_{\rm CC}\ \Ycc\\
        \alpha_{\rm CC} &= \frac{\Yct - \alpha_{\rm AGB}\ \Ycagb}{\Ycc}
    \end{align}
\end{subequations}
Above, $\alpha_{\rm CC}$ is dependent on the choice of $\alpha_{\rm AGB}$ to keep the total C yield constant.

In Section \ref{sec:f-z-models}, we find that $\alpha_{\rm AGB}\approx 2.9$ for the \cxi{} yield model. Using this \agb\ yield, we can now use Eq.~\ref{eq:y_c_eq} to find an estimate of $\Ycc$ as a function of metallicity. We show the values of $\Ycc$ we obtain from the subgiant sample in Fig.~\ref{fig:analytic}. From a regression analysis, we suggest here that
\begin{equation}\label{eq:zeta}
    \Ycc = 0.0028 + \zeta\left(\frac{Z}{Z_\odot}\right),
\end{equation}
where $\zeta\approx0.001$ describes the metallicity dependence.

@@

Classical, \textit{one-zone} models of chemical evolution assume instantaneous mixing of metals in the star-forming interstellar medium \citep[e.g.][]{matteucci21}. This simple framework is a poor approximation of the Milky Way.  The Galaxy evolves insideout---where star formation is higher towards the center and in the early universe \citep{bird+13}. Additionally, stars can migrate several kpcs over their lifetimes, mixing different chemical environments across the galaxy \citep{bird+12,sellwood+binney02}. For the rest of this paper, we focus on multi-zone models, which discretize the Galaxy into concentric rings in which stars move between.  
Our models extends the \citet{james+21} (hereafter \JJ) Milky Way model, run with the publicly available Versatile Integrator for Chemical Evolution (\VICE). 
This model is described extensively in \JJ~and concisely summarized  in \citet{james+23}. Here, we provide a brief overview of the relevant model components.

>>>>

% citations
\defcitealias{james+21}{J21}
\newcommand{\JJ}{\citetalias{james+21}}
\newcommand{\VICE}{\textsc{vice}}
\newcommand{\citetjack}{Roberts et al.~(2023, in prep.)}
\newcommand{\citepjack}{(Roberts et al.~2023, in prep.)}
\newcommand{\citealtjack}{Robert et al.~2023, in prep.}
\newcommand{\cxi}{\texttt{\hyperlink{C11}{C11}}}
\newcommand{\kx}{\texttt{\hyperlink{K10}{K10}}}
\newcommand{\kxvi}{\texttt{\hyperlink{K16}{K16}}}
\newcommand{\vxiii}{\texttt{\hyperlink{V13}{V13}}}

% Acronyms
\newcommand{\cc}{CCSNe}
\newcommand{\agb}{AGB}
\newcommand{\ia}{SNe Ia}
\newcommand{\sfh}{SFH}
\newcommand{\dla}{DLA}
\newcommand{\ssp}{SSP}
\newcommand{\imf}{IMF}
\newcommand{\gce}{GCE}
\newcommand{\apogee}{APOGEE}

% internal abbreviations
\newcommand{\lims}{lower-intermediate mass stars}
\newcommand{\hms}{high-mass stars}
\newcommand{\caah}{[C/Mg]-[Mg/H]}
\newcommand{\caafe}{[C/Mg]-[Mg/Fe]}

% other
\makeatletter
\newcommand{\C}[1][\@nil]{
    \def\tmp{#1}%
    \ifx\tmp\@nnil%
        \ensuremath{\rm C}%
    \else%
        \ifmmode ^{#1}{\rm C}%
        \else $^{#1}$C%
        \fi%
\fi }
\makeatother

\newcommand{\proton}{\isotope[1][1]{p}}
\newcommand{\Yct}{\ensuremath{y_{\rm C}}}
\newcommand{\Ycc}{\ensuremath{y_{\rm C}^{\rm cc}}}
\newcommand{\Yoc}{\ensuremath{y_{\rm Mg}^{\rm cc}}}
\newcommand{\Ycagb}{\ensuremath{y_{\rm C}^{\rm agb}}}

    \newcommand{\y}{\ensuremath{\rotatebox[origin=B,y=0.5ex]{180}{y}}}
\newcommand{\Mo}{%
    \ifmmode {\rm M}_{\sun}%
    \else M$_{\sun}$
    \fi}
\newcommand{\Zo}{%
    \ifmmode Z_{\sun}%
    \else $Z_{\sun}$%
    \fi}

\newcommand{\about}[1]{${\sim} #1$}

@@

To quantify yields, we define the stellar yield to be the fraction of a stars initial mass which is newly synthesized and released as a given element. 
For an element $X$ and star with mass $M$, the net-fractional stellar yield $\y$ is 
\begin{equation}
\y_{X} = \frac{M_{\rm ejected}}{M_{\rm birth}} \Delta Z_X
\end{equation}
where $M_{\rm ejected}$ and $M_{\rm birth}$  are the total ejected mass and the birth mass of the star, and $\Delta Z_X$ is the change in mass fraction of abundance X from birth to ejected. For example, if a 1\,\Mo\ star has $\y_{\C} = 10^{-3}$, then the star will add $10^{-3}\,\Mo$ of new C to the interstellar medium. 
Although per-star yields are necessary to compute \agb{} star enrichment rates in \gce{}  models, \imf-averaged yields are useful in interpreting their predictions. 
Given the fractional yield $\y(M, Z)$ as a function of initial stellar mass $M$ and metallicity $Z$, the \imf-averaged yield is given by 
\begin{equation} \label{eq:imf-yield}
    y_{\rm X}(Z,t) = 
    \frac{
    \int_{M_{\rm min}(t)}^{M_{\rm max}} 
    \y_{\rm X}(M, Z)
    \frac{dN}{dM}\ M\ dM
}
{
    \int \frac{dN}{dM}\ M\ dM
}
\end{equation}
where ${dN}/{dM}$ is the \imf, $M_{\rm max}=100\,\Mo$ is the maximum stellar mass, and $M_{\rm min}(t)$ is the mass of stars with lifetime $t$.\footnotemark{}
We use $t=10\,$Gyr for total yields when $t$ is not used.
To calculate the \imf-averaged net yields, we use the Versatile Integrator for Chemical Evolution code (\VICE\footnotemark{}).

@@

	\begin{tabular}{l l l l}
		\hline
        Element & $\Ycc$ & $\y^{\rm AGB}$ & $y^{\rm SNe Ia}$ \\
		\hline
        C & $0.0028 + 0.001\,(Z/Z_\odot)$ & $2.9\times{\rm C11}$ &  0 \\
        Mg & 0.00185 & 0 & 0 \\
        Fe & 0.0012 & 0 & 0.00214 \\
        N & 0.00072 & $9\times10^{-4}\,(Z/Z_\odot)\,M$ & 0\\
		\hline
	\end{tabular}
\end{table}

@@

An \agb\ star is a low-mass star ($\lesssim 8 M_{\sun}$) during its final phase of evolution.  
In an \agb\ star, two competing processes determine the outcome of C production: \textit{third dredge up} and \textit{hot bottom burning}.  
Third dredge up accompanies thermal pulses in \agb\ stars, where material from the CO core is mixed with the envelope, increasing surface C abundances \citep{KL14}. The C yields of the star are increased as this C-enhanced envelope is released to the interstellar medium. 
Hot bottom burning is the activation of the CNO cycle\footnotemark{}
at the bottom of the convective envelope when $T\gtrsim 50\,{\rm MK}$. Because the $^{14}$N proton capture is the slowest component of the CNO cycle, the CNO cycle converts nearly all \C[12] into $^{14}$N \citep{solar-fusion}.

\footnotetext{
    The CNO cycle is a series of proton-capture reactions with CNO elements resulting in energy generation and the creation of an $\alpha$ particle. $\C[12]({\rm p}, \gamma)
    ^{13}{\rm N}(\beta^+, \nu_{\rm e})
    ^{13}{\rm C}({\rm p}, \gamma)\allowbreak
    ^{14}{\rm N}({\rm p}, \gamma)\allowbreak
    ^{15}{\rm O}(\beta^+, \nu_{\rm e})\allowbreak
    ^{15}{\rm N}({\rm p}, \alpha)
    \C[12]$. 
There are other less important minor branches of the CNO cycle
 \citep{solar-fusion}.
}

Hot bottom burning and third dredge-up result in mass-dependent C yields. 
Stars less than about??2\,\Mo do not experience third dredge-up. As a result, these stars C abundances are only affected by first dredge-up, resulting in little change to C yields or slight destruction of C.
In 2?? to 4?? \Mo stars, third dredge up becomes important, enriching the outer layers with C, making these stars the most important producers of C. 
In massive, high-metallicity \agb\ stars, both hot bottom burning and third dredge up occur; however, third dredge up is much more efficient, resulting in significant \C[12] destruction.

    In this work, we explore four different sets of \agb\ star yield tables from literature,
    providing necessary well-sampled grids in mass and metallicity. We refer to the yields from the following studies as the following.
\begin{itemize}
    \item \cxi: \citet{cristallo+11, cristallo+15}
    \item \kx: \citet{karakas10}
    \item \vxiii: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \kxvi: \citet{KL16, karakas+18}
\end{itemize}
In Appendix~\ref{sec:oob_models}, the masses, metallicites, and notable differences between each yield set are briefly described. 
For our models to match observations, we find that need to uniformly amplify these yield tables. We use \cxi\ table, amplified by a factor of 2.9, as the fiducial \agb\ yield.

@@

Fig.~\ref{fig:agb-ssp}, on the left, shows the total production of C by \agb\ stars in a \ssp{} at an age $t$, i.e. $\y_{\rm C}(Z_\odot, t)$. 
As the mass range $2\,\Mo\lesssim M \lesssim 4\,\Mo$ is most important for C production, about half of C production occurs before \about{1}\,Gyr, similar to \ia\ Fe. 
\kx{} and \kxvi{} weight C production more heavily towards high-mass AGB stars resulting in a faster enrichment delay time, whereas the \cxi\ and \vxiii\ models predict a slightly longer timescale of \about{1}\,Gyr. In any case, little to no C is produced more than 2\,Gyr after a star formation event. In contrast, Fe production continues steadily for 10\,Gyr.

@@

\begin{figure*}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}
        \caption[]{The net fractional \agb\ C yield  plotted as a function of initial stellar mass $M$ and colour-coded according to metallicity. The black dashed line shows $\y=0$ for reference. Each panel represents yields from one of four \agb\ models: \cxi, \kx{}, \vxiii{}, \kxvi{} (see Section \ref{sec:agb} and  Appendix \ref{sec:oob_models}). }
        \label{fig:y_agb}
\end{figure*}

@@

Massive stars form $^{12}$C in their cores through the triple--$\alpha$ process. However, only C ejected through supernovae and stellar winds contributes to the yield. 
While there are many stellar models providing predictions of \cc{} yields, the results of these models are highly uncertain due to the many stellar modeling uncertainties.

@@

\cc\ models do not reach [O/Mg] due to overproduction of O or underproduction of Mg, or both \citep{emily+21}. Here, we assume ${\rm [O/Mg]} = 0$, which is not compatible with \cc\ models but is consistent with \apogee{} observations \citep{weinberg+19, weinberg+22}.

Out model predictions of [C/Mg] and [C/H] are identical. There is a slight slope in [O/Mg] with metallicity, so this could add a systematic to our model.

As we focus on constraining relative yields, we neglect O and Mg yield variations in the main text (excluding the uniform scaling of yields and massloading in Section~\ref{sec:outflows}). There is substantial variation in predicted Mg yields. Most models predict relatively flat trends metallicity (even with rotation as in \citealt{LC18}). However, the variation is significant and our adopted $\Yoc$ yield is much higher than most models. This is a known problem (see \citealt{emily+21}). \cc\ models underpredict $[{\rm Mg}/{\rm O}]$, and the reason why is unknown.

@@

We assume a simple \textit{one-zone} chemical evolution model \cite[e.g.][]{tinsley80, pagel09, matteucci21}.  Newly produced metals are homogeneously and instantaneously mixed, so spatial dependence is neglected.
We define $M_{X}$ to be the mass of element $X$ in the gas-phase, $\dot{M}_\star$ to be the star formation rate (in $\Mo\,{\rm yr}^{-1}$), and $\eta$ to be the mass loading factor $\eta\equiv\dot{M}_{\rm outflow}/\dot{M}_\star$ (representing the strength of outflows). A \ssp\ returns a fraction $r$ of their birth mass to the interstellar medium, due to ejected stellar envelopes.\footnotemark{} 
Here, we will use $\alpha$ as a representative $\alpha$-element, such as O or Mg. 
Given the \imf-averaged yield $y_\alpha$, the rate of change in the gas-phase mass of $\alpha$ is a simple sum of sources and sinks,
\begin{equation}
    \dot{M}_\alpha =  y_\alpha\,\dot{M}_\star - \dot{M}_{\alpha,\,\text{remnants}} - \dot{M}_{\alpha,\,\text{outflows}}
\end{equation}
where $y_\alpha\,\dot{M}_{\star}$ describes \cc\ enrichment. 
In terms of the return mass fraction of stars $r$, the mass lost to remnants is $Z_\alpha\,\dot{M}_\star\,(1-r)$.  And, the outflows deplete mass at a rate $Z_\alpha \,\dot{M}_\star\,\eta$. (We assume the composition of outflows is the same as the interstellar medium.) Substituting for $\eta$ and $r$,  
\begin{equation}
    \dot{M}_\alpha= y_\alpha\,\dot{M}_\star - (1 + \eta - r)\,Z_\alpha\,\dot{M}_\star.
\end{equation}
Assuming an exponentially declining star formation history $\dot{M}_\star \propto e^{-t/\tau_{\rm sfh}}$, the equilibrium abundance is derived analytically by setting $\dot{Z}_\alpha=0$.
\begin{equation}\label{eq:z_eq}
    Z_\alpha^{\rm eq}(R) = \frac{y_\alpha}{1 + \eta(R) - r - \tau_\star / \tau_{\rm sfh}},
\end{equation}
where $\tau_{\star}$ is the star formation rate. ($\eta$ depends on $R$ to create a metallicity gradient. See also Section \ref{sec:vice} and Eq.~\ref{eq:eta_r}.)
In the special case of constant star formation, $\tau_{\rm sfh}\to\infty$, the denominator simplifies to $1+\eta-r$.

Likewise, for \agb\ stars,%
\footnote{In detail, the \agb\ contribution requires integration over the \sfh\ to account for the finite lifetimes of stars. However, for this exponential \sfh, this effect is minimal, so we simply use $\Ycagb$ for the current \agb\ contribution.}
Eq.~\ref{eq:z_eq} then becomes
\begin{equation}
    Z_{\rm C}^{\rm eq}(R) = \frac{\Ycc + \Ycagb}{1 + \eta(R) - r - \tau_\star / \tau_{\rm sfh}}
\end{equation}
And the equilibrium C/$\alpha$ abundance ratio is
\begin{equation}\label{eq:z_co}
    \frac{Z_{\rm C}^{\rm eq}}{Z_\alpha^{\rm eq}} = \frac{\Ycc + \Ycagb }{y_\alpha}.
\end{equation}
Analogous to \cite{james+23} arguments about N, the trends in abundance ratios are set by yield ratios in these \gce{} models. The effect of other \gce{} parameters (most importantly $\eta$) cancels. As a consequence, yield ratios should establish abundance ratio trends in models which assume a different normalization of element yields and mass-loading (see discussion below).
This argument can also be inverted to infer  yields from abundance ratio trends. To the extent that observed C and $\alpha$ trends reflect the equilibrium abundances in different Galactic regions, we can infer the \cc\ yield given an assumed \agb\ star yield (or vice versa). Inferring $\Ycc$ from $\Ycagb$, 
\begin{equation}
    y_\text{C}^\alpha =  y_\alpha^\text{CC} \frac{Z_\text{C,\,eq}}{Z_{\alpha,\,{\rm eq}}} - \Ycagb
\end{equation}
Rewriting this expression as a relative yield of C and Mg,
\begin{equation}\label{eq:y_c_eq}
    \frac{\Ycc}{y_\alpha} = \left(\frac{Z_{\text{C}}}{Z_{\alpha}}\right)_\odot 10^{[{\rm C}/\alpha]} - \frac{\Ycagb}{y_\alpha}.
\end{equation}

@@

As we will discuss in Section \ref{sec:outflows}, the normalization of yields and $\eta$ is degenerate. This can be observed in Eq.~\ref{eq:z_co}, where changes in $\eta$ or the scaling of $y_{\rm C}/y_{\rm Mg}$ would not affect equilibrium trends. Furthermore, in Eq.~\ref{eq:z_eq}, an increase in both $y_{\rm Mg}$ and $\eta$ would leave $Z_{\rm Mg}^{\rm eq}$ unchanged. Our models here are unable to distinguish the overall scaling of yields and outflow mass loading. Choosing to keep $y_{\rm C}$ fixed reduces unnecessary free parameters. So, at solar metallicity, we set
\begin{subequations}
    \begin{align}
        \Yct(\Zo) &= \Ycc + \Ycagb = 0.005 \\
        \left(\frac{\Yct}{\Yoc}\right)_{\Zo} &= 2.7.
    \end{align}
\end{subequations}
This ratio results in an equilibrium abundance ${\rm [C/Mg]} = -0.09$, which is consistent with the subgiant sample and is within \about{20\%} of the solar C/Mg mixture from \citet{asplund+09}.

In Section \ref{sec:f-z-models}, we will show that none of the four \agb\ yield sets (\cxi{}, \kx{}, \vxiii{}, \kxvi{}) produce enough C relative to this $\Ycc$ value. So, we introduce normalization factors, $\alpha_{\rm AGB}$, and $\alpha_{\rm CC}$ which denote a multiplicative scaling of $\Ycc$ and $\Ycagb$ respectively. 
\begin{subequations} \label{eq:alpha}
    \begin{align}
        \Ycagb &\rightarrow \alpha_{\rm agb}\ \Ycagb \\
        \Ycc &\rightarrow \alpha_{\rm cc}\ \Ycc\\
        \alpha_{\rm CC} &= \frac{\Yct - \alpha_{\rm agb}\ \Ycagb}{\Ycc}
    \end{align}
\end{subequations}
Above, $\alpha_{\rm cc}$ is dependent on the choice of $\alpha_{\rm agb}$ to keep the total C yield constant.

In Section \ref{sec:f-z-models}, we find that $\alpha_{\rm agb}\approx 2.9$ for the \cxi{} yield model. Using this \agb\ yield, we can now use Eq.~\ref{eq:y_c_eq} to find an estimate of $\Ycc$ as a function of metallicity. We show the values of $\Ycc$ we obtain from the subgiant sample in Fig.~\ref{fig:analytic}. From a regression analysis, we suggest here that
\begin{equation}\label{eq:zeta}
    \Ycc = 0.0028 + \zeta\left(\frac{Z}{Z_\odot}\right),
\end{equation}
where $\zeta\approx0.001$ describes the metallicity dependence.

@@

Classical, \textit{one-zone} models of chemical evolution assume instantaneous mixing of metals in the star-forming interstellar medium \citep[e.g.][]{matteucci21}. This simple framework is a poor approximation of the Milky Way.  The Galaxy evolves \textit{inside-out}---where star formation is higher towards the center and in the early universe \citep{bird+13}. Additionally, stars can migrate several kpc over their lifetimes, mixing different chemical environments across the galaxy \citep{bird+12,sellwood+binney02}. For the rest of this paper, we focus on multi-zone models, which discretize the Galaxy into concentric rings in which stars move between.  
Our models extends the \citet[hereafter \JJ]{james+21} Milky Way model, run with the publicly available Versatile Integrator for Chemical Evolution (\VICE). 
This model is described extensively in \JJ~and concisely summarized  in \citet{james+23}. Here, we provide a brief overview of the relevant model components.
